### 2014-05-12 paper/main.tex
Add beginnings of ACL-style paper

%
% File acl2014.tex
%
% Contact: koller@ling.uni-potsdam.de, yusuke@nii.ac.jp
%%
%% Based on the style files for ACL-2013, which were, in turn,
%% Based on the style files for ACL-2012, which were, in turn,
%% based on the style files for ACL-2011, which were, in turn, 
%% based on the style files for ACL-2010, which were, in turn, 
%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
%% based on the style files for EACL-2009 and IJCNLP-2008...

%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{acl2014}
\usepackage{times}
\usepackage{microtype}
\usepackage{url}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{paralist} % inline lists
\usepackage{commands}

% smart quotes
\usepackage[autostyle, english=american]{csquotes}
\MakeOuterQuote{"}

%\setlength\titlebox{5cm}

% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\title{Deep Neural Models for Bilingual Lexicon Extraction}

\author{Jon Gauthier \\
  Symbolic Systems Program \\
  Stanford University \\
  Stanford, CA 94305 \\
  {\tt jgauthie@stanford.edu} \\\And
  Arthur Tsang \\
  Department of Computer Science \\
  Stanford University \\
  Stanford, CA 94305 \\
  {\tt atsang2@stanford.edu} \\}

\date{}

\begin{document}
\maketitle

\begin{abstract}
  We present a new method for producing bilingual lexicons from very
  non-parallel corpora. In contrast with previous approaches, our method
  separates the steps of \begin{inparaenum}[(1)]
    \item constructing distributed word representations from monolingual
      corpora, and
    \item building a mapping between these representations.
  \end{inparaenum}

  % TODO
  We demonstrate that a deep neural model can effectively learn a translational
  mapping between two monolingual vector spaces for the purposes of bilingual
  lexicon extraction. We evaluate the performance of this model relative to X
  and Y.
\end{abstract}

\section{Introduction}

In the task of bilingual lexicon extraction (BLE), we accept as input some pair
of corpora in different languages---perhaps parallel or perhaps not---and learn
associations of translational equivalence between one language (a "source
language") and the other (a "target language").

Bilingual lexicon extraction is obviously of most utility when applied to
under-resourced language pairs which do not already benefit from a surplus of
lexicon materials. These same language pairs, lacking resources as simple as
translation dictionaries, also lack sufficient aligned text for training
BLE models. For this reason, most recent work on this task focuses on
minimally supervised approaches which do not require parallel corpora
\cite{rapp1995,peirsman2010}.

\newcite{rapp1995} suggested that there is "a correlation between the patterns
of word co-occurrences in different languages." In line with this hypothesis,
recent work has determined the likelihood of two words as translations of one
another by comparing the translations of the contexts in which they appear. The
standard approach is to manually\unclear construct a "bilingual vector space" which
contains distributed word representations of words in both the source and target
language \cite{fung1998,peirsman2010,vulic2013}.

\section*{Acknowledgments}

The acknowledgments should go immediately before the references.  Do
not number the acknowledgments section. Do not include this section
when submitting your paper for review.

\bibliographystyle{acl}
\bibliography{bibliography}

\end{document}