### 2014-06-07 paper/main.tex
Clean up paper LaTeX, begin extending for long paper version

%
% File acl2014.tex
%
% Contact: koller@ling.uni-potsdam.de, yusuke@nii.ac.jp
%%
%% Based on the style files for ACL-2013, which were, in turn,
%% Based on the style files for ACL-2012, which were, in turn,
%% based on the style files for ACL-2011, which were, in turn,
%% based on the style files for ACL-2010, which were, in turn,
%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
%% based on the style files for EACL-2009 and IJCNLP-2008...

%% Based on the style files for EACL 2006 by
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{acl2014}
\usepackage{times}
\usepackage{microtype}
\usepackage{url}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{paralist} % inline lists
\usepackage{commands}
\usepackage{cleveref}
\usepackage{placeins} % FloatBarrier

\usepackage{tikz}
\usepackage{pgfplots}

% DEV
\usepackage{booktabs}

% smart quotes
\usepackage[autostyle, english=american]{csquotes}
\MakeOuterQuote{"}

%\setlength\titlebox{5cm}

% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\title{Deep Neural Networks for Bilingual Lexicon Extraction}

\author{Jon Gauthier,\affilA{} Arthur Tsang,\affilB{} Christopher Potts\affilC{} \\
  \affilA{} Symbolic Systems Program, \affilB{}Department of Computer Science, \affilC{}Department of Linguistics \\
Stanford University, Stanford, CA 94035 \\
{\tt \{jgauthie,atsang2,cgpotts\}@stanford.edu}}

\date{}

\begin{document}
\maketitle

\begin{abstract}
  We explore a method for producing bilingual lexicons from non-parallel
  corpora. In contrast with most previous approaches, our method separates the
  steps of \begin{inparaenum}[(1)]
  \item constructing distributed word representations from monolingual corpora,
    and
  \item building a translational mapping.
  \end{inparaenum}

  This gives us precise control over the independent development of the word
  representations and the mapping between them.

  We demonstrate that a deep neural network can effectively learn a
  translational mapping between these representations for the purposes of
  bilingual lexicon extraction more accurately than a closely related linear
  model.
\end{abstract}

\section{Introduction}
\label{sec:introduction}

In bilingual lexicon extraction (BLE), we accept as input some pair of corpora
in different languages---perhaps parallel or perhaps not---and use them to learn
associations of translational equivalence between a \textit{source language} and
a \textit{target language}.

Bilingual lexicon extraction is of most utility when applied to under-resourced
language pairs which do not already benefit from a surplus of lexicon materials.
These same language pairs, lacking resources as simple as quality translation
dictionaries, also lack sufficient amounts of manually aligned text to train BLE
models. For this reason, most recent work on this task focuses on minimally
supervised approaches which do not require parallel corpora
\cite{rapp1995,peirsman2010}.

\newcite{rapp1995} suggested that there is "a correlation between the patterns
of word co-occurrences in different languages." In line with this hypothesis,
recent work has related the likelihood of two words being translations of one
another to the similarities between the contexts in which the words appear in
their corresponding languages. The standard approach under this theory is to
manually construct a \textit{bilingual vector space} which contains distributed
representations of words in both the source and target language
\cite{fung1998,vulic2013}. The axes of this space are built from
bilingual seed pairs. A given word representation has a large value along an axis
if it co-occurs often with the seed word corresponding to that axis. Because
axes are defined with bilingual pairs, word representations for both source and
target language words can lie in the same space. With a bilingual vector space
constructed, translations for a given source-language word can be retrieved by
simply finding the nearest target-language word representation neighboring the
word's own vector representation in the bilingual vector space.

The present paper is motivated by the hypothesis that this standard bilingual
vector space approach to BLE is inherently limited by the design of the space
itself. We suggest that the construction of the space along axes composed of a
hand-picked set of bilingual seeds has the potential to limit the expressiveness
of the word representations contained therein. Consider, for example, a vector
space constructed in this fashion using just two seeds: \textit{hot} and
\textit{cold}. Such a space may be effective at representing the difference
between words like \textit{fire} and \textit{ice}. However, with these two axes
it would fail to capture many other distinctions, such as the difference between
the words \textit{up} and \textit{down}. The quality of the vector space is thus
heavily dependent on the quality of its seeds.

\newcite{mikolov2013exploiting} suggest an alternate method for bilingual
lexicon extraction, centered around two monolingual vector spaces $E$
(corresponding to the source language) and $F$ (corresponding to the target
language). These vector spaces are constructed using a standard neural language
model~\cite{mikolov2013}. This is a promising alternative to the aforementioned
bilingual vector space approach, as it allows for the construction of robust
distributed representations whose quality is not dependent on manually selected
seeds. The authors find that a linear model trained via stochastic gradient
descent to map between these two vector spaces can effectively find word-level
translations from the source to the target language. We extend this work in
attempting to approximate a \emph{translation function} $t : E \to F$ using a
deep neural network, trained to accept as input word vectors from the source
language space $E$ and output word vectors in the target language space $F$. We
capitalize on the fact that these vector spaces may be independently
constructed, experimenting with different configurations for each space.

% \Cref{sec:data} describes the corpora used to test this translation process.
% \Cref{sec:model} details the actual process of translation. In
% \Cref{sec:evaluation,sec:conclusion} we analyze the performance of our model and
% examine how it supports our motivating hypothesis.

\section{Data}
\label{sec:data}

Our translation algorithm makes use of two monolingual corpora: one consisting
of text in the source language and the other containing text in the target
language. We require only that the corpora have the following properties:

\subparagraph{Large and broad-domain.} With this property given, we can safely
assume that for any common word which appears in the source-language corpus, we
will also find its translation somewhere in the target-language corpus. If large
corpora are unavailable, we can make use of smaller narrow-domain corpora (e.g.,
religious texts, famous novels).

\subparagraph{Non-parallel.} This means that our algorithm does not depend on
any sort of alignment, lexical or topical. This relative flexibility allows us
to demonstrate that the method can be applied in situations where parallel
corpora are not available.

% TODO for long paper: describe preprocessing techniques (stolen from Gensim).
In our experiments we use English and Spanish corpora sourced from Wikipedia.%
\footnote{Available at \url{http://dumps.wikimedia.org/}.} We choose not to take
advantage of the document alignments available in the Wikipedia corpus in order
to show that our method is generalizable to other corpora.

In addition to the two monolingual corpora, we train on a minimal set of word
translation pairs. The specifics of these training pairs are described in
\Cref{subsec:translation-nn}.

\section{Model}
\label{sec:model}

\begin{figure}[tb!]
	\centering
	\resizebox{\linewidth}{!}{
    \begin{tikzpicture}
        \node (english) at (0, 2) {\large \textbf{book}};

        \draw (3.25, 2) ellipse[x radius=0.75, y radius=1.1];
        \node (englishvec) at (3.25, 2) {$\begin{bmatrix}~\vdots~\\\vdots\end{bmatrix}$};

        \draw (6.5, 2) ellipse[x radius=0.75, y radius=1.1];
        \node (spanishvec) at (6.5, 2) {$\begin{bmatrix}~\vdots~\\\vdots\end{bmatrix}$};

        \node (spanish) at (9.5, 2) {\large \textbf{libro}};

        \draw [->] (english) -- (englishvec) node [pos=0.45, above] {\small \texttt{word2vec}};
        \draw [->] (englishvec) -- (spanishvec) node [midway, above, align=center] {Neural\\network};
        \draw [->] (spanishvec) -- (spanish) node [pos=0.55, above, align=center] {Nearest\\neighbor};
    \end{tikzpicture}
    }
    \caption{The process of word translation, using a neural network to map
      between two monolingual VSMs}
    \label{fig:translation}
\end{figure}

\Cref{fig:translation} illustrates the process of translating a single word from
a source language to a target language (here English and Spanish, respectively)
in this model. We begin with an input source-language word and retrieve a
corresponding pre-computed word representation in the source VSM\@. We use this
source-language word vector as an input to a neural network, which outputs a
vector in the target-language VSM\@. We perform a simple nearest-neighbor search
in the target-language VSM to find the closest existing word vector, and take
its corresponding target-language word to be the translation of the original
source-language word.

\subsection{Word representations}
\label{subsec:word-representations}

For each of the two corpora introduced in \Cref{sec:data}, we train a continuous
vector space model on 10-word contexts, using an implementation in the
\textit{Gensim} software package of \texttt{word2vec}, a neural language
model~\cite{rehurek2010,mikolov2013}. This neural language model attempts to
learn word representations which minimize error in the task of predicting a
word's context given the word itself. The resulting word vectors are thus
theoretically optimal (i.e., minimum-cost) representations of the meanings
underlying their corresponding words. This is in contrast with the word
representations built by a standard BLE approach, which define themselves in
terms of manually constructed axes, and thus do not necessarily account for a
maximal amount of variance in meaning among the words represented. We experiment
with different word vector dimensionalities, and detail our specific choices
later in \Cref{subsec:test-data}.

While we focus on constructing a neural language model with \texttt{word2vec},
we believe any distributional model of word meaning would suffice for this
task~\cite{bengio2003,turian2010,collobert2011}.

\subsection{Translation with a neural network}
\label{subsec:translation-nn}

The central computation is the transformation of a source-language word
representation into a target-language word representation via a deep neural
network. We learn this transformation from a small set of training pairs.

The standard bilingual vector space approach detailed in the introduction would
base its word representations on these training pairs. In contrast with this
method, our model uses these translation pairs only in its final learning step,
while training the neural network which models the translation function. For
this reason, the model can benefit from effective word representations developed
independently from the translation system itself. Furthermore, it can provide
direct feedback through a training cost function on the quality of the
translation pairs.

We investigate several different methods of producing translation pairs. In line
with previous approaches, we fill a set of translation pairs \texttt{Cog} with
cognates shared between the two languages~\cite{koehn2002}. We also build a
small separate set \texttt{ConcNN} consisting of concrete nouns, and attempt to
restrict the set to word pairs for which both the source- and target-language
words have just one sense in their respective languages.

For each training pair of words $(e, f)$, we provide as a training example the
input and output $(w_e, w_f)$, where $w_e$ and $w_f$ are word representations
drawn from their respective VSMs constructed as described in
\Cref{subsec:word-representations}.

At inference time, for a given English word $e$, we provide as input to the
neural network the corresponding word representation $w_e$. The output of the
network, $w_f$, is some vector in the target-language vector space. We search
for the nearest known neighbor of $w_f$, and treat the word corresponding to
this nearest neighbor as the translation $f$ of the input word $e$.

\subsection{Comparison: Translation with a linear transformation}
\label{subsec:translation-linear}

\newcite{mikolov2013exploiting} found that a linear model could learn a
translation mapping between two monolingual vector spaces. We construct such a
linear model for purposes of comparison with the neural network method detailed
above. It utilizes the same basic idea as presented in \Cref{fig:translation}:
it only differs in the type of transformation used to map between the two
monolingual vector spaces. Whereas our approach uses a deep neural network, this
linear model builds a linear transformation between the two spaces. Given
source-language and target-language training data matrices $W_E, W_F$ (where the
$i$th columns of the two matrices are the vector word representations
corresponding to the $i$th training pair) we assert the existence of a
translation matrix $A$, where $W_F = AW_E$.

We compute the linear transformation by taking $A = W_F W_E^+$, where $W_E^+$ is
the Moore-Penrose pseudoinverse of $W_E$. It is necessary to use the
pseudoinverse of $W_E$, as the matrix is unlikely to be square and invertible.
Then given any source-language word vector $w_e$, we calculate the transformed
target-language word vector as $w_f = A w_e$.

\section{Evaluation}
\label{sec:evaluation}
We evaluate our translation model on an unseen set of English-Spanish word
pairs. These test sets are customized to match the algorithm's training
environment. For example, a translation algorithm trained on concrete nouns from
the training set \texttt{ConcNN} will be tested on other unseen concrete nouns.
Since our model builds a ranked list of nearest-neighbor translation candidates,
we can calculate the mean reciprocal rank (MRR) of the correct translation
within the list:
\begin{equation}
    \textit{MRR} = \frac{1}{n}\sum_{j=1}^n \frac{1}{r_j}
\end{equation}
where $r_j$ is the rank of the correct translation for pair $j$ out of $n$. For
example, a mean reciprocal rank of 1 would indicate that every translation
output ranked the correct translation in the first position. A mean reciprocal
rank of $\frac{1}{2}$ would indicate that the average translation output placed
the correct translation as the second most likely in a ranked list.

\subsection{Baseline}
\label{subsec:baseline}

We establish a lower bound for translation performance with an algorithm that
associates source- and target-language words by their relative corpus
frequencies. Given an input source-language word, we compute its corpus
frequency percentile. We then predict translations in the target language by
returning the words with the closest percentile frequencies in the target
language corpus.

\subsection{Test data}
\label{subsec:test-data}

We establish three separate test environments, each defined by a choice of
training data, input corpus format, and VSM design. The vector space
dimensionalities given below were derived through trial and error. We find that
different VSM dimensionalities are optimal for different corpus inputs. Overall,
we see that performance varies significantly based on these test environment
properties. The environments are as follows:
\begin{description}
\item[\texttt{Cog}:] Minimally preprocessed corpora, \texttt{Cog} training data,
  500-dimensional word vectors % TODO long paper: 5000/750 min word threshold
\item[\texttt{ConcNN}:] Minimally preprocessed corpora, \texttt{ConcNN} training
  data, 500-dimensional word
  vectors % TODO long paper: 10000/1500 min word threshold
\item[\texttt{CogLem}:] Lemmatized Spanish corpus, \texttt{Cog} training data,
  1000-dimensional word vectors % TODO long paper: 2500/375 min word threshold
\end{description}

\begin{table*}[htb!]
    \centering
    \small
    \begin{tabular}[t]{l lll}
    \toprule
    Diagnosis & Input & Expected cognate output & Predicted ranked output (translations) \\
    \midrule
    Synonyms & \emph{acceptance} & \emph{aceptación} & \emph{aprobación (approval), \dots} \\
    Near misses & \emph{vanilla} & \emph{vainilla} & \emph{chocolate, vainilla, \dots} \\
    Right neighborhood, wrong spot & \emph{election} & \emph{elección} & \emph{candidato, candidatura (candidacy), \dots} \\
    Alternate sense interpretations & \emph{division} & \emph{división} & \emph{unidad (unit), élite, comando (command), \dots} \\
    \bottomrule
    \end{tabular}
    \caption{Examples of common error patterns observed in the algorithm's output for
        the \bt{Cog} environment}
    \label{tbl:errors}
\end{table*}

Our evaluation metric looks for a single possible translation of a test input
word in ranked output. As we explain in \Cref{subsec:error-analysis}, this may
needlessly penalize the algorithm, as it often suggests reasonable translations
earlier in the ranked output which happen to not match our strict search
requirements. For example, our algorithm is tested on a cognate pair
$(\textit{concept}, \textit{concepto})$, and is penalized for suggesting the
translation \textit{idea}, a suitable synonym of the desired target word. A
fourth test run \bt{CogFix} considers the algorithm's output on the \texttt{Cog}
dataset, and accepts such sensible alternate translations (rather than
stipulating in this case that all source-language cognates translate to their
corresponding target-language cognates).

\subsection{Results}
\label{subsec:results}

\begin{table}[tb]
    \centering
%    \begin{subtable}%{0.44\linewidth}
        \resizebox{\linewidth}{!}{
        \begin{tabular}[t]{l cccc}
            % CogFix: manual fixes (allow plurals, synonyms, etc.)
            \toprule
            Model & \bt{Cog} & \bt{ConcNN} & \bt{CogLem} & \bt{CogFix} \\
            \midrule
            NN (125) & 0.216 & 0.199 & 0.215 & 0.389 \\
            NN (250) & \textbf{0.295} & 0.193 & 0.269 & \textbf{0.480} \\
            NN (500) & 0.264 & \textbf{0.203} & 0.279 & 0.432 \\
            NN (1000) & 0.263 & 0.196 & 0.265 & 0.451 \\[1.5ex]
            Linear & 0.280 & 0.195 & \textbf{0.286} & 0.449 \\
            Baseline & 0.006 & 0.002 & 0.002 & 0.006 \\
            \bottomrule
        \end{tabular}
        }
%    \end{subtable}
    \caption{Performance on multiple test environments (hidden layer dimensionalities given in parentheses). MRR values (higher is better).}
    \label{tbl:results}
%    \hfill
%    \begin{subfigure}{0.55\linewidth}
%        \resizebox{\linewidth}{!}{
%        \begin{tikzpicture}
%            \begin{axis}[xlabel=Hidden layer dimensionality, ylabel=MRR, legend pos=outer north east,
%              legend style={font=\small}]
%                \addplot[red, mark=x] coordinates {(125, 0.2158) (250, 0.2945) (500, 0.2635) (1000, 0.2628)};
%                \addlegendentry{Cog}
%                \addplot[red, dashed, forget plot, domain=125:1000] {0.2803};
%
%                \addplot[gray, mark=square*] coordinates {(125, 0.1991) (250, 0.1933) (500, 0.2029) (1000, 0.1956)};
%                \addlegendentry{ConcNN}
%                \addplot[gray, dashed, forget plot, domain=125:1000] {0.1954};
%
%                \addplot[blue, mark=*] coordinates {(125, 0.2147) (250, 0.2694) (500, 0.2788) (1000, 0.2646)};
%                \addlegendentry{CogLem}
%                \addplot[blue, dashed, forget plot, domain=125:1000] {0.2861};
%
%                \addplot[brown, mark=triangle*] coordinates {(125, 0.3891) (250, 0.4795) (500, 0.4318) (1000, 0.4509)};
%                \addlegendentry{CogFix}
%                \addplot[brown, dashed, forget plot, domain=125:1000] {0.4494};
%          \end{axis}
%        \end{tikzpicture}
%        }
%        \subcaption{Mean reciprocal ranks for various training environments (higher is better). Baseline performance for each seed set is illustrated as a flat dashed line in matching color.}
%    \end{subfigure}
%    \caption{Evaluation results}
%    \label{fig:results}
\end{table}

Mean reciprocal rank values for each test environment and a number of training
conditions are in \Cref{tbl:results}. The linear transformation model based on
\newcite{mikolov2013exploiting} was very competitive with the neural network
model. This suggests that many---but not all---of the geometric relationships in
the target language vector space can be modeled through a simple linear
transform of the geometric structures in the source language space.

The best absolute performance is achieved by a neural network with 250
hidden-layer neurons in the \bt{CogFix} training environment. Interestingly, the
hidden layer of this model is of lower dimensionality than the test input and
output word vectors (each 500 dimensions). This suggests that the model learns
some useful compressed intermediate representation of the input source-language
word in the translation training process. Its MRR shows us that on average the
algorithm suggests an acceptable translation in the second or third position of
the ranked list of possible translations.

\subsection{Error analysis}
\label{subsec:error-analysis}

While the translation algorithm fails to predict the exactly correct translation a
significant amount of the time, an analysis of common prediction errors is encouraging.
We identify several common error patterns in the
algorithm's output in \Cref{tbl:errors}, and elaborate on the classes of errors below:

\subparagraph{Synonyms.} With many of the "errors" of this form, the algorithm
provided an acceptable alternate translation that we simply did not expect.
These cases are treated as successes in the \bt{CogFix} training environment, as
described in \Cref{subsec:results}.

\subparagraph{Near misses.} In these cases the algorithm mapped into a precise
neighborhood of related words, but picked the wrong element from the set. For
example, the output word \textit{chocolate} is strongly related to the expected
word \textit{vainilla}. % TODO long paper: future work might examine
% the actual distance values and "give up" if the top few words are all very
% close

\subparagraph{Right neighborhood, wrong spot.} We occasionally mapped into a
correct larger area of the target language vector space, but still made a
significant semantic error. With the input word \textit{election}, for example,
the algorithm predicts as translations target-language words related to
political campaigns.

\subparagraph{Alternate senses.} In several cases on the \texttt{Cog} training
set and its derivatives, the algorithm learned to predict translations of the
input words which corresponded to valid alternate interpretations of the source
language word. For example, it interprets the English \textit{division} in the
sense of a military unit (rather than in the mathematical sense as we expected).
These cases were not treated as successes in the \bt{CogFix} training
environment. We hope to eliminate this class of errors in future work by using
multi-prototype VSMs (see \Cref{sec:conclusion}).

% TODO long paper: Show T-SNE visualization of the neighborhoods into which we map erroneously --- show that we still map into the correct neighborhood, just the wrong spot, a lot of the time

\section{Conclusion}
\label{sec:conclusion}

In this paper, we presented an algorithm which learns word translations from
unlabeled, non-parallel corpora by building a mapping between two independently
constructed monolingual vector spaces. This algorithm can be used to perform
bilingual lexicon extraction in scenarios where the source and target language
in question do not benefit from widely available parallel corpora. We
demonstrated encouraging translation performance with large, broad-domain
corpora as text sources.

For simplicity, our current VSM design associates a single word vector with each
word in the source or target language corpus. However, training error could be
greatly reduced by allowing polysemous words to be associated with multiple word
vectors~\cite{reisinger2010,huang2012}. We believe such an extension will remove
much of the remaining ambiguity in the input vector spaces.

% However, as we note that many of our errors are caused by alternate sense interpretations, our approach could be further improved through disambiguation and multiple vector representations per word.

\bibliographystyle{acl}
\bibliography{bibliography}

\end{document}